### 2014-10-28 thesis-writing/intro.tex
Fixed some typos, added some todo's.

\section{Introduction}
Garbage collection (GC) is very useful for software development. It helps the developers to avoid explict deallocation and also avoid dangling pointers. There has been constant complaint that the GC reduces the perfomance of the application. But the recent research shows that there is infact much benefit in using it. With the advent of multicore, the processor utilization is of the high concern and application post multi-core era has been focusing on building concurrent applications. Garbage Collections are designed to be concurrent and parallel. The distintion between concurrent and parallel is if the application runs simultaneously with the garbage collection it is termed as concurrent collector while there are more than one collector thread runs during the collection, it is called as parallel collector. There has been lot of focus on concurrent collectors lately.

In HPC community, applications are mostly written in languages with no or less managed runtime systems. It is just time that all applications that are written for the huge clusters to change to managed runtime systems. The HPC applications can benefit all the advantages of the garbage collection. Predominantly HPC codes are written for the two kind of the memory architectures. The first being the shared memory systems, where the memory is common for all the processing element. Other being distributed memory system where each processing element has its own share of memory. Bacon et al described the theory of garbage collection in very lucid way. The proposed hyphothesis is based on the Bacon et al observation on the different kinds of garbage collector. The two main GC techniques are mark-sweep (MS) and reference counting (RC). All the available GC algorithm follows one of these approach to identify garbage. In the abstract way, MS traces all the live heap objects and then deletes the untraced but allocated heap objects. On the other hand, RC traces all the dead/potential dead objects and then deletes the traced objects.

The hyphothesis of this research is from the above abstarct idea. If MS collector can determine the garbage by traversing the live heap objects, then in a huge memory systems with more live objects the collector should perform worse than RC collectors. In reality, MS is most preffered GC than RC. It is constantly proved that RC performs worse than MS. RC implementations have write barriers which require additional instruction to update the reference count of the heap objects on every change in the object graph.
In a highly optimized GC system with huge memory, RC based systems will perform better than MS. The research steer towards RC based garbage collection.

>>>>

\section{Introduction}
Garbage collection (GC) is very useful for software development. It helps the developers to avoid explicit deallocation and also avoid dangling pointers. There has been constant complaint that the GC reduces the performance of the application. But the recent research shows that there is in fact much benefit in using it. With the advent of multicore, the processor utilization is of the high concern and application post multi-core era has been focusing on building concurrent applications. Garbage Collections are designed to be concurrent and parallel. The distinction between concurrent and parallel is if the application runs simultaneously with the garbage collection it is termed as concurrent collector while there are more than one collector thread runs during the collection, it is called as parallel collector. There has been lot of focus on concurrent collectors lately.

In HPC community, applications are mostly written in languages with no or less managed runtime systems. It is just time that all applications that are written for the huge clusters to change to managed runtime systems. The HPC applications can benefit all the advantages of the garbage collection. Predominantly HPC codes are written for the two kind of the memory architectures. The first being the shared memory systems, where the memory is common for all the processing element. Other being distributed memory system where each processing element has its own share of memory. Bacon et al described the theory of garbage collection in very lucid way. The proposed hypothesis is based on the Bacon et al observation on the different kinds of garbage collector. The two main GC techniques are mark-sweep (MS) and reference counting (RC). All the available GC algorithm follows one of these approach to identify garbage. In the abstract way, MS traces all the live heap objects and then deletes the untraced but allocated heap objects. On the other hand, RC traces all the dead/potential dead objects and then deletes the traced objects.

The hypothesis of this research is from the above abstract idea. If MS collector can determine the garbage by traversing the live heap objects, then in a huge memory systems with more live objects the collector should perform worse than RC collectors. In reality, MS is most preferred GC than RC. It is constantly proved that RC performs worse than MS. RC implementations have write barriers which require additional instruction to update the reference count of the heap objects on every change in the object graph.
In a highly optimized GC system with huge memory, RC based systems will perform better than MS. The research steer towards RC based garbage collection.

### 2014-10-28 thesis-writing/litrev.tex
Fixed some typos, added some todo's.

\section{Literature Review}
McCarthy invented the concept of Garbage Collection in 1959 ~\cite{mccarthy}. His works focused on indirect method of collecting 
garbage which is now called mark-sweep or tracing collector. While Mark-sweep was the only method that was available at the early 1960's,
 Collins designed a new method called reference counting, which is more direct method to detect garbage~\cite{collins}. 
 Dijkstra et al~\cite{dijkstra} describes the correctness condition of tricolor tracing based collector. This work redefined the 
 way the collectors should be designed and how one can prove the properties of the collectors. Haddon et al ~\cite{haddon} realized 
 the fragmentation issues in the mark-sweep algorithm and designed a mark-compact algorithm. Fenichel et al~\cite{feni} and Cheney~\cite{cheney}
 provided new kind of algorithm based on copying the live objects which was called copying colletors. Based on Foderaro and Fateman ~\cite{fode81} observation, 
 98\% of the objects collection has been allocated after the last collection. Several such observations were made and reported in ~\cite{zorn89,sans93}.
 Appel designed a simple generation garbage collection~\cite{Appel89}.

 While most of the literature up until now considered a indirect method of collecting garbage through tracing, the other form of collection is counting the references.
 While counting the references seems obvious way to solve the problem, McBeth's~\cite{McBeth1983} article on the inability to collect cyclic garbage structures by 
 reference counting method made the method less practical.

Bobrow's ~\cite{Bobrow1980} solution to collect cylic garbage based on reference counting method was the first hybrid collector which uses both tracing and
counting the references. Hughes ~\cite{hugh83,hugh87} identified the issues in the Bobrow's algorithm.
It was proven by McBeth \cite{McBeth1963} in early sixties that reference counting collectors were unable to handle cyclic structures; several attempts to fix this problem appeared subsequently, e.g. \cite{Friedman1979,Bobrow1980,Lins2008}. We give details in Section \ref{subsection:otherrelated}.
In contrast to the approach followed in \cite{Friedman1979,Bobrow1980,Lins2008} and several others,
Brownbridge \cite{Brownbridge1985} proposed, in 1985, a strong/weak pointer algorithm to tackle the problem of reclaiming cyclic data structures by distinguishing cycle closing pointers (weak pointers) from other references (strong pointers) \cite{Jones1996}. %different algorithm to tackle the problem of reclaiming cyclic structures using strong and weak pointers.
This algorithm relied on maintaining two invariants: (a) there are no cycles in strong pointers and (b) all items in the graph must be strongly reachable from the roots.

@@

% Note: Gokarna had a lot of this backwards. He was contrasting reference counting
% to stop-the-world collectors, and the kinds of write barriers he was concerned with
% are present in our system.
With the advancement of multiprocessor architectures,
reference counting garbage collectors have become popular because
they do not require all application threads to be stopped before the garbage collection algorithm can run~\cite{Levanoni2006}.
Recent work in reference counting algorithms, e.g. \cite{Barabash2005,Levanoni2006,Bacon2001,Bacon:2001:JWC}, try to
reduce concurrent operations and increase the efficiency of reference counting collectors.
Since our collector is a reference counting collector, it can potentially benefit from the
same types of optimizations discussed here. We leave that, however, to a future work.

>>>>

\section{Literature Review}
\todo{...on an indirect method...now called a mark-sweep...}
McCarthy invented the concept of Garbage Collection in 1959 ~\cite{mccarthy}. His works focused on indirect method of collecting 
garbage which is now called mark-sweep or tracing collector. While Mark-sweep \todo{You didn't capitalize last time} was the only method that was available at the early 1960's, \todo{This first clause does nothing for the sentence. You could probably omit it.}
 Collins designed a new method called reference counting, which is more direct method to detect garbage~\cite{Collins1960}. 
 Dijkstra et al~\cite{dijkstra} describes the correctness condition of tricolor tracing based collector. This work redefined the 
 way the collectors should be designed and how one can prove the properties of the collectors. Haddon et al ~\cite{haddon} realized 
 the fragmentation issues in the mark-sweep algorithm and designed a mark-compact algorithm. Fenichel et al~\cite{feni} and Cheney~\cite{cheney}
 provided new kind of algorithm based on copying the live objects which was called copying collectors. Based on Foderaro and Fateman ~\cite{fode81} observation, 
 98\% of the objects collection has been allocated after the last collection. Several such observations were made and reported in ~\cite{zorn89,sans93}.
 Appel \todo{Apple?} designed a simple generation garbage collection~\cite{Appel89}.

 While most of the literature up until now considered a indirect method of collecting garbage through tracing, the other form of collection is counting the references.
 While counting the references seems obvious way to solve the problem, McBeth's~\cite{McBeth1963} article on the inability to collect cyclic garbage structures by 
 reference counting method made the method less practical.

Bobrow's ~\cite{Bobrow1980} solution to collect cyclic garbage based on reference counting method was the first hybrid collector which uses both tracing and
counting the references. Hughes ~\cite{hugh83,hugh87} identified the issues in the Bobrow's algorithm.
It was proven by McBeth \cite{McBeth1963} in early sixties that reference counting collectors were unable to handle cyclic structures; several attempts to fix this problem appeared subsequently, e.g. \cite{Friedman1979,Bobrow1980,Lins2008}. We give details in Section \ref{subsection:otherrelated}.\todo{This reference doesn't work}
In contrast to the approach followed in \cite{Friedman1979,Bobrow1980,Lins2008} and several others,
Brownbridge \cite{Brownbridge1985} proposed, in 1985, a strong/weak pointer algorithm to tackle the problem of reclaiming cyclic data structures by distinguishing cycle closing pointers (weak pointers) from other references (strong pointers) \cite{Jones1996}. %different algorithm to tackle the problem of reclaiming cyclic structures using strong and weak pointers.
This algorithm relied on maintaining two invariants: (a) there are no cycles in strong pointers and (b) all items in the graph must be strongly reachable from the roots.

@@

% Note: Gokarna had a lot of this backwards. He was contrasting reference counting
% to stop-the-world collectors, and the kinds of write barriers he was concerned with
% are present in our system.
With the advancement of multiprocessor architectures,
reference counting garbage collectors have become popular because
they do not require all application threads to be stopped before the garbage collection algorithm can run~\cite{Levanoni2006}.
Recent work in reference counting algorithms, e.g. \cite{Barabash2005,Levanoni2006,Bacon2001,Bacon:2001:JWC}, try to
reduce concurrent operations and increase the efficiency of reference counting collectors.
\todo{Forward reference. You haven't described our collector.}
Since our collector is a reference counting collector, it can potentially benefit from the
same types of optimizations discussed here. We leave that, however, to a future work.
